Morgue: Supplementary material

### 2018-06-08 morgue_supp_mat.Rmd
adding some things I don't anticipate needing anymore

ggplot_uniFL <- function(FL.df) {
  ggplot(FL.df, aes(x = Var1, y = mean_fitness)) +
    geom_line() + 
    geom_ribbon(aes(ymin = lower_2.5, ymax = upper_97.5), alpha = 0.5)
}

>>>>

```{r gsg Comparison}

## Fit generalized additive mixed models ----
gamm_control_2 <- gamm.model(gall_survival ~ s(sc.gall_size) + s(sc.clutch_size,k=9) + s(sc.female_preference),
                             data = control_df)

gamm_treatment_2 <- gamm.model(gall_survival ~ s(sc.gall_size) + s(sc.clutch_size,k=9) + s(sc.female_preference),
                             data = treatment_df)

## Estimate directional, nonlinear, and correlational selection gradients ----
sg.size_control <- gradient.calc(mod = gamm_control_2$gam, 
                                 phenotype = "sc.gall_size", 
                                 covariates = c("sc.clutch_size","sc.female_preference"))
sg.size_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                   phenotype = "sc.gall_size", 
                                   covariates = c("sc.clutch_size","sc.female_preference"))
sg.clutch_control <- gradient.calc(mod = gamm_control_2$gam, 
                                        phenotype = "sc.clutch_size", 
                                        covariates = c("sc.gall_size","sc.female_preference"))
sg.clutch_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                          phenotype = "sc.clutch_size", 
                                          covariates = c("sc.gall_size","sc.female_preference"))
sg.preference_control <- gradient.calc(mod = gamm_control_2$gam, 
                                    phenotype = "sc.female_preference", 
                                    covariates = c("sc.clutch_size","sc.gall_size"))
sg.preference_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
                                      phenotype = "sc.female_preference", 
                                      covariates = c("sc.clutch_size","sc.gall_size"))
#sg.size.x.clutch_control <- gradient.calc(mod = gamm_control_2$gam, 
#                                               phenotype = c("sc.gall_size", "sc.clutch_size"),
#                                               covariates = c("sc.female_preference"))
#sg.size.x.clutch_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
#                                                 phenotype = c("sc.gall_size", "sc.clutch_size"),
#                                                 covariates = c("sc.female_preference"))
#sg.size.x.preference_control <- gradient.calc(mod = gamm_control_2$gam, 
#                                               phenotype = c("sc.gall_size", "sc.female_preference"),
#                                               covariates = c("sc.clutch_size"))
#sg.size.x.preference_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
#                                                 phenotype = c("sc.gall_size","sc.female_preference"),
#                                                 covariates = c("sc.clutch_size"))
#sg.preference.x.clutch_control <- gradient.calc(mod = gamm_control_2$gam, 
#                                                  phenotype = c("sc.female_preference", "sc.clutch_size"),
#                                                  covariates = c("sc.gall_size"))
#sg.preference.x.clutch_treatment <- gradient.calc(mod = gamm_treatment_2$gam, 
#                                                    phenotype = c("sc.female_preference", "sc.clutch_size"),
#                                                    covariates = c("sc.gall_size"))

## Summarise the results ----
#sg.summary <- bind_rows(
#          data.frame(sg.size_control$ests,  
#                     gradients = rownames(sg.size_control$ests),
#                     treatment = "Complex"),
#          data.frame(sg.size_treatment$ests, 
#                     gradients = rownames(sg.size_treatment$ests),
#                     treatment = "Simple"),
#          data.frame(sg.clutch_control$ests,  
#                     gradients = rownames(sg.clutch_control$ests),
#                     treatment = "Complex"),
#          data.frame(sg.clutch_treatment$ests,  
#                     gradients = rownames(sg.clutch_treatment$ests),
#                     treatment = "Simple"),
#          data.frame(sg.preference_control$ests, 
#                     gradients = rownames(sg.preference_control$ests),
#                     treatment = "Complex"),
#          data.frame(sg.preference_treatment$ests, 
#                     gradients = rownames(sg.preference_treatment$ests),
#                     treatment = "Simple"))#,
          #data.frame(sg.size.x.clutch_control$ests,  
          #           gradients = rownames(sg.size.x.clutch_control$ests),
          #           treatment = "Complex")[5,],
          #data.frame(sg.size.x.clutch_treatment$ests,  
          #           gradients = rownames(sg.size.x.clutch_treatment$ests),
          #           treatment = "Simple")[5,],
          #data.frame(sg.size.x.preference_control$ests,  
          #           gradients = rownames(sg.size.x.preference_control$ests),
          #           treatment = "Complex")[5,],
          #data.frame(sg.size.x.preference_treatment$ests,  
          #           gradients = rownames(sg.size.x.preference_treatment$ests),
          #           treatment = "Simple")[5,],
          #data.frame(sg.preference.x.clutch_control$ests, 
          #           gradients = rownames(sg.preference.x.clutch_control$ests),
          #           treatment = "Complex")[5,],
          #data.frame(sg.preference.x.clutch_treatment$ests, 
          #           gradients = rownames(sg.preference.x.clutch_treatment$ests),
          #           treatment = "Simple")[5,])

## Table of Estimates ----
#sg.table <- sg.summary %>%
#  mutate(plus_minus = "\u00B1", estimates = round(estimates, 4), SE = round(SE, 4)) %>%
#  unite(col = estimate_SE, from = c("estimates", "plus_minus", "SE"), sep="") %>%
#  select(-P.value) %>%
#  spread(treatment, estimate_SE)
#sg.table
#write_csv(sg.table, "sg.table.csv")

## Table of P-values ----
#sg.summary %>%
#  select(-estimates, -SE) %>%
#  spread(treatment, P.value)

## BOOTSTRAPPED INTERVALS ----
bootstrap_df <- bind_rows(
  data.frame(Bootstrap_Estimates = sg.size_control$boot[ ,1], Phenotype = "Gall size", Selection_Type = "Beta", Food_Web = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size_control$boot[ ,2], Phenotype = "Gall size", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.size_treatment$boot[ ,1], Phenotype = "Gall size", Selection_Type = "Beta", Food_Web = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.size_treatment$boot[ ,2], Phenotype = "Gall size", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.clutch_control$boot[ ,1], Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.clutch_control$boot[ ,2], Phenotype = "Clutch size", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.clutch_treatment$boot[ ,1], Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.clutch_treatment$boot[ ,2], Phenotype = "Clutch size", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.preference_control$boot[ ,1], Phenotype = "Female preference", Selection_Type = "Beta", Food_Web = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.preference_control$boot[ ,2], Phenotype = "Oviposition preference", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.preference_treatment$boot[ ,1], Phenotype = "Female preference", Selection_Type = "Beta", Food_Web = "Simple"))#,
#  data.frame(Bootstrap_Estimates = sg.preference_treatment$boot[ ,2], Phenotype = "Oviposition preference", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  # correlational
#  data.frame(Bootstrap_Estimates = sg.size.x.clutch_control$boot[ ,5], Phenotype = "Size,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size.x.clutch_treatment$boot[ ,5], Phenotype = "Size,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Simple"),
#    data.frame(Bootstrap_Estimates = sg.size.x.preference_control$boot[ ,5], Phenotype = "Size,Pref.", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size.x.preference_treatment$boot[ ,5], Phenotype = "Size,Pref.", Selection_Type = "Correlational", Food_Web_Treatment = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.preference.x.clutch_control$boot[ ,5], Phenotype = "Pref.,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.preference.x.clutch_treatment$boot[ ,5], Phenotype = "Pref.,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Simple")
#)

bootstrap_df %>%
  #filter(Selection_Type == "Gamma") %>%
  ggplot(., aes(x=Phenotype, y=Bootstrap_Estimates, fill=Food_Web)) +
  geom_boxplot() +
  geom_hline(yintercept=0, linetype="dotted") +
  facet_wrap(~Selection_Type, scales = "free_x") +
  coord_flip()

gsg_gradients <- bootstrap_df %>%
  group_by(Phenotype, Food_Web) %>%
  summarise(Beta = round(mean(Bootstrap_Estimates),3),
            lower_2.5 = round(quantile(Bootstrap_Estimates, probs = 0.025),3),
            upper_97.5 = round(quantile(Bootstrap_Estimates, probs = 0.975),3)) %>%
  mutate(Model = "gsg_GAMM")

bind_rows(gsg_gradients, summary_Beta_glmm_control, summary_Beta_glmm_treatment) %>%
  arrange(Phenotype, Food_Web, Model) %>%
  mutate(absolute_range = upper_97.5 - lower_2.5) %>%
  knitr::kable(.)
```

```{r} 
ggplot_uniFL <- function(FL.df) {
  ggplot(FL.df, aes(x = Var1, y = mean_fitness)) +
    geom_line() + 
    geom_ribbon(aes(ymin = lower_2.5, ymax = upper_97.5), alpha = 0.5)
}
```

```{r Selection Gradients - Bootstrapped Intervals, eval=FALSE}

## BOOTSTRAPPED INTERVALS ----
bootstrap_df <- bind_rows(
  data.frame(Bootstrap_Estimates = sg.size_control$boot[ ,1], Phenotype = "Gall size", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size_control$boot[ ,2], Phenotype = "Gall size", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.size_treatment$boot[ ,1], Phenotype = "Gall size", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.size_treatment$boot[ ,2], Phenotype = "Gall size", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.clutch_control$boot[ ,1], Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.clutch_control$boot[ ,2], Phenotype = "Clutch size", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.clutch_treatment$boot[ ,1], Phenotype = "Clutch size", Selection_Type = "Beta", Food_Web_Treatment = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.clutch_treatment$boot[ ,2], Phenotype = "Clutch size", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  data.frame(Bootstrap_Estimates = sg.preference_control$boot[ ,1], Phenotype = "Female preference", Selection_Type = "Beta", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.preference_control$boot[ ,2], Phenotype = "Oviposition preference", Selection_Type = "Gamma", Food_Web_Treatment = "Complex"),
  data.frame(Bootstrap_Estimates = sg.preference_treatment$boot[ ,1], Phenotype = "Female preference", Selection_Type = "Beta", Food_Web_Treatment = "Simple"))#,
#  data.frame(Bootstrap_Estimates = sg.preference_treatment$boot[ ,2], Phenotype = "Oviposition preference", Selection_Type = "Gamma", Food_Web_Treatment = "Simple"),
  # correlational
#  data.frame(Bootstrap_Estimates = sg.size.x.clutch_control$boot[ ,5], Phenotype = "Size,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size.x.clutch_treatment$boot[ ,5], Phenotype = "Size,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Simple"),
#    data.frame(Bootstrap_Estimates = sg.size.x.preference_control$boot[ ,5], Phenotype = "Size,Pref.", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.size.x.preference_treatment$boot[ ,5], Phenotype = "Size,Pref.", Selection_Type = "Correlational", Food_Web_Treatment = "Simple"),
#  data.frame(Bootstrap_Estimates = sg.preference.x.clutch_control$boot[ ,5], Phenotype = "Pref.,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Complex"),
#  data.frame(Bootstrap_Estimates = sg.preference.x.clutch_treatment$boot[ ,5], Phenotype = "Pref.,Clutch", Selection_Type = "Correlational", Food_Web_Treatment = "Simple")
#)

bootstrap_df %>%
  #filter(Selection_Type == "Gamma") %>%
  ggplot(., aes(x=Phenotype, y=Bootstrap_Estimates, fill=Food_Web_Treatment)) +
  geom_boxplot() +
  geom_hline(yintercept=0, linetype="dotted") +
  facet_wrap(~Selection_Type, scales = "free_x") +
  coord_flip()

gsg_gradients <- bootstrap_df %>%
  group_by(Phenotype, Food_Web_Treatment) %>%
  summarise(Beta = round(mean(Bootstrap_Estimates),3),
            lower_2.5 = round(quantile(Bootstrap_Estimates, probs = 0.025),3),
            upper_97.5 = round(quantile(Bootstrap_Estimates, probs = 0.975),3))

knitr::kable(gsg_gradients)
```

```{r}
library(schoenberg)
pointwise <- confint(gamm_red.control$gam, parm = "Gall_Height_mm", trans = function(x) {exp(x)/(1+exp(x))})
simultaneous <- confint(gamm_red.control$gam, parm = "Gall_Height_mm", trans = function(x) {exp(x)/(1+exp(x))}, type = "simultaneous", seWithMean=TRUE)

compare.confint <- bind_rows(mutate(pointwise, confint = "pointwise"), mutate(simultaneous, confint = "simultaneous")) 
compare.confint$confint <- relevel(factor(compare.confint$confint), ref = "simultaneous")

ggplot(compare.confint, aes(x=Gall_Height_mm)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.25) +
  geom_line(aes(y=est), color="black") +
  facet_wrap(~confint)

schluter <- gam.plot(gamm_red.control$gam)
schluter

gam.plot(glmm_lin.red.control$gam, all.terms=TRUE)
test <- plot(glmm_lin.red.control$gam, all.terms=TRUE)
gam.gradients(mod = glmm_lin.red.control$gam, phenotype = "Gall_Height_mm", se.method = "n")
coef(glmm_lin.red.control$gam)

# test derivative calculations
control_df$Genotype <- factor(control_df$Genotype)
control_df$Plant_Position <- factor(control_df$Plant_Position)
control_df$Gall_Number <- factor(control_df$Gall_Number)
A <- gam(gall_survival ~ s(Gall_Height_mm) + s(Plant_Position, bs="re") + s(Gall_Number, bs="re"), data = control_df, family=binomial(link=logit)) # random=~(1|Genotype/Plant_Position/Gall_Number),
summary(A)

head(control_df)
A_withMean <- plot(A, se = TRUE, seWithMean = T, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))}); A_withMean
A_only <- plot(A, se = TRUE, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))}); A_only
draw(A)

# new data for prediction
newDF <- with(control_df, data.frame(Gall_Height_mm = seq(min(Gall_Height_mm),max(Gall_Height_mm),length.out = 100)))
# prediction of smoothed estimates at each unique year value
# with standard error    
B <- predict(A$gam,  newDF, type="response", se.fit=TRUE)
B

# finite difference approach to derivatives following
# example from ?predict.gam

eps <- 1e-7
X0 <- predict(A, newDF, type = 'lpmatrix')
X0 <- exp(X0)/(1+exp(X0))

newDFeps_p <- newDF + eps

X1 <- predict(A, newDFeps_p, type = 'lpmatrix')
X1 <- exp(X1)/(1+exp(X1))
# finite difference approximation of first derivative
# the design matrix
Xp <- (X0 - X1) / eps
Xp
# first derivative
fd_d1 <- Xp %*% coef(A)
fd_d1/mean_fit
# second derivative
newDFeps_m <- newDF - eps

X_1 <- predict(A$gam, newDFeps_m, type = 'lpmatrix')
# design matrix for second derivative
Xpp <- (X1 + X_1 - 2*X0)  / eps^2
# second derivative
fd_d2 <- Xpp %*% coef(A$gam)

exp(fd_d1) / ( 1+exp(fd_d1))  # type = "response"
exp(fd_d2) / ( 1+exp(fd_d2))  # type = "response"

mean(fd_d1)/mean_fit
mean(fd_d2)/mean_fit

gam.gradients(A, phenotype = "Gall_Height_mm", se.method="n", standardized = FALSE)

coef(A)
mean_fit <- mean(control_df$gall_survival)
fderiv_size <- fderiv(A, newdata = data.frame(Gall_Height_mm = mean(control_df$Gall_Height_mm)), offset = mean_fit)
test <- fderiv_size$derivatives$`s(Gall_Height_mm)`$deriv
exp(test)/(1+exp(test))

plot(A, se = TRUE, seWithMean = T, shift = mean(predict(A)), trans = function(x) {exp(x)/(1+exp(x))})
diff <- .Machine$double.eps^(1/3)
predicts <- predict(A, newdata = data.frame(Gall_Height_mm = c(mean(control_df$Gall_Height_mm)-diff,mean(control_df$Gall_Height_mm)+diff)), type="lpmatrix")
tests <- predicts %*% coef(A)
exp(tests)/(1+exp(tests))
t1 <- exp(predicts[2])/(1+exp(predicts[2]));t1
t2 <- exp(predicts[1])/(1+exp(predicts[1]));t2
((t1-t2)/(diff*2))/mean(control_df$gall_survival)
gam.gradients(A, phenotype = "Gall_Height_mm", se.method="n", standardized = FALSE)

	Wbar <- function(x, mod, phenotype,covariates) {
    		new.d <- as.data.frame(mod$model[,c(phenotype,covariates)])
    		names(new.d)<-c(phenotype,covariates)
    		new.d2<-new.d
    		for (i in 1:length(x)) {
    		    new.d[,as.character(phenotype[i])] <-
    		    new.d[,as.character(phenotype[i])]+x[i]
    		}

    		p <- predict.gam(
			   object=mod,
    	 	   newdata=new.d,
    	 	   newdata.guaranteed=TRUE,
    	 	   type="response"
    		)
    	return(mean(p))
	}
	as.data.frame(A$model[ ,"Gall_Height_mm"])
	#mean(control_df$gall_survival)
	Wbar(x=0, mod=A, phenotype="Gall_Height_mm", covariates=NULL)
first.derivatives <- grad(f=Wbar, x=rep(0, 1), mod=A, phenotype="Gall_Height_mm", covariates=NULL); first.derivatives
denom <- Wbar(x=rep(0,1), mod=A, phenotype="Gall_Height_mm", covariates=NULL); denom
first.derivatives/denom
pracma:::grad
A$Vp
A$coefficients
A_withMean
```
